Comparing R packages for calculation of persistent homology

### 2020-02-20 tdabench-paper/tdabench.Rmd
edited narrative for fig cir

\subsubsection{Runtime as a function of data dimension (feature = n - 1)}
For small point cloud numbers, Gudhi and TDAstats performed similarly while Dionysus lagged behind. 
At larger point clouds, TDAstats outperforms Gudhi, and the difference among the Rips libraries is more visible.
This result is especially true for calculating higher feature dimensions on large point clouds. 
While the runtime values changed for each point cloud condition, the growth pattern with respect to increasing number of points was similar for all Rips complex libraries. 
For the conditions tested in our analysis, all Rips libraries resulted in an apparent polynomial growth for run time. 
The exact values for the log-log regression of this growth pattern is provided in supplemental data on the paper Github.

Each panel in figure \ref{fig:cir} corresponds to an n-dimensional circle (circle, sphere, and hypersphere). 
For each point cloud, n-1 dimensional features were calculated.
As the sphere dimensions and feature dimensions increased, the polynomial time growth becomes apparent sooner. 
At these higher dimensional features, some of the curves seem to cut off. 
This is due to certain point cloud sizes not being computable on our devices. 
These limitations are laid out in the "Computation Limits" section.

>>>>

Tori do not trivially generalize to dimensions other than 3, but circles do.
Benchmarking on a circular point cloud permits confirmation of the pattern in Figure \ref{fig:tor} while also revealing how the libraries compare as dataset dimension increases.
Figure \ref{fig:cir} exhibits the resulting data for a 2-dimensional circle (panel A), 3-dimensional sphere (panel B), and 4-dimensional hypersphere (panel C).
When the dataset dimension equals 2, GUDHI practically matches Ripser's performance in outpacing Dionysus.
However, in the case of the 3-dimensional sphere, the pattern visualized in Figure \ref{fig:tor} for the 3-dimensional torus is again present.
By the 4th dimension, the gap between Ripser and GUDHI widens.
Of note, missing points for larger datasets in Figure \ref{fig:cir} are not plotted if and only if calculating persistent homology caused an error due to insufficient RAM.
Thus, for the hypersphere, Ripser was able to calculate persistent homology for a dataset with approximately 3 times as many points as Dionysus and over 2 times as many as GUDHI.
Interestingly, all curves plotted in Figure \ref{fig:cir} grow polynomially with respect to the number of points (see Supplement for regression details).

For each point cloud, n-1 dimensional features were calculated. *previous sentence will be moved to figure caption*
